\chapter{Introduction \& Motivation}
\label{chap:intro}

\Acp{vpn} play an important role in todays network infrastructure. %TODO reference
They are used in a multitude of scenarios and by very different actors. Multi-national firms use them to inter-connect their spread locations across the world. 
So the usage ranges from many hundreds (or thousand) of concurrent sessions to just one in the private user setup.

- mobile telephone base station to backbones 
Goal: connect and transmit data over the public Internet

With these many different usage scenarios, a variety of protocols and implementations were developed over the years, each with slightly different opinions about many aspects.

As the amounts of traffic increases over time \cite{DECIX} the implementations handling it must become more efficient at it.
%TODO: Comparison CPU speed vs. traffic increase (vs. encryption speeds?) over time

Specialized hardware exists, but is expensive. So low-cost software solutions with off-the-shelf hardware remain relevant.

\section{Research question}
This thesis develops a performance model for \Acp{vpn} which categorizes and ranks different possible performance hazards in protocols and implementations. Additionally it will show which bottlenecks are worth optimizing compared to the effort required and the expectable gains.

This model is applied to common software \Acp{vpn} implementations.

We analyze how and why common bottlenecks occur, and present approaches to solve them.

Lastly we investigate if our implementation of the WireGuard protocol, called MoonWire, based on the learned facts is faster than the canonical version and how much.

\section{Structure of Thesis}
The remainder of this thesis is structured as follows.
Chapter~\ref{chap:background} gives an overview on the technical background surrounding VPN technology, introduces the necessary cryptographic foundations and presents various protocols and implementations. 
Related work in the area of VPN performance measurements can be found in Chapter~\ref{chap:related_work}. %TODO
In Chapter~\ref{chap:perf_model} we develop the performance model and list which factors it includes.
The design and implementation details of MoonWire are explained in Chapter~\ref{chap:moonwire}, where dive into the technical details of approaches, explain their trade-offs and present performance figures.
In Chapter~\ref{chap:evaluation} we benchmark MoonWire against the developed performance model and compare the results to the other VPN implementations, before we conclude in Chapter~\ref{chap:conclusion}.

